### 2022-08-19 thesis/appendix.tex
Formatting, font changes. Rearrangment of figures

\appendix

\chapter{Additional Benchmark Model Definitions}
\label{app:benchmark}

@@

\noindent The model was first studiend in \cite{fitzhugh1961impulses} and its reachability analysis using Bernstein expansion is studied in \cite{dang2012reachability}.

>>>>

\begin{appendices}

\chapter{ADDITIONAL BENCHMARK MODEL DEFINITIONS}
\label{app:benchmark}

@@

\noindent The model was first studied in \cite{fitzhugh1961impulses} and its reachability analysis using Bernstein expansion is considered in \cite{dang2012reachability}.

\end{appendices}

### 2022-08-19 thesis/bern_prune_wrapping_error.tex
Formatting, font changes. Rearrangment of figures

Difference between Queried Min Coeff and Actual Min Coeff: -
0.719813449727491
\end{verbatim}

\caption{Example of Wrapping Error Obstacle demonstrated for Step 4 for Vanderpol Model.}
\label{fig:bern_prune_wrapping_output}
\end{figure}

>>>>

      Difference between Queried Min Coeff and Actual Min Coeff:
      -0.719813449727491
\end{verbatim}

\caption{Example of Wrapping Error Obstacle Demonstrated for Step 4 for Vanderpol Model.}
\label{fig:bern_prune_wrapping_output}
\end{figure}

### 2022-08-19 thesis/formats.tex
Formatting, font changes. Rearrangment of figures

\section{Drawbacks to the Static Algorithm}
Before we embark on introducing the dynamic algorithm, let us first consider a few motivating factors.
%
As discussed in Section \ref{sec:static}, the static algorithm requires template directions to be specified by the user in order for the reachable set computation to proceed.
%
The parallelotopes tended to be defined by a combination of axis-aligned directions or diagonal directions.
%
However, it is unclear whether these template directions work reasonably well over general non-linear dynamics.
%
This is a problem which has profound consequences for computing useful over-approximations.
%
Choosing the incorrect directions could very well lead to an over-approximation too conservative for any practical use.
%
Figure \ref{fig:bloat_cvdp} portrays a simple example of a reachable set whose wrapping error becomes explosive after a few steps.
%
In light of these ramifications, it is esstentially incumbent on the user to choose fruitful template directions that control the growth of the over-approximation.
%
Due to the inherent difficult nature of choosing these template directions, it may be unreasonable to delegate this responsibility to a practitioner who knows little about parallelotope-based reachability.
%
These drawbacks to the static algorithm motivate the rationale behind the \emph{automatic} aspect of our modified algorithm.

Additionally, as the template directions are set at the beginning before the reachable set computation proceeds, these directions cannot adapt to the dynamics as the computation executes.
%
There may be template directions which may yield much leaner over-approximations at an intermediate step during the computation.
%
Hence, the ability to change the composition of the parallelotope bundle by adding and removing parallelotopes based on a data-driven approach could yield leaner over-approximations.
%
These considerations motivate the rationale behind the \emph{dynamic} aspect of our modified algorithm.

>>>>

\section{Drawbacks to the Static Algorithm}
Before we embark on introducing the dynamic algorithm, let us first consider a few motivating factors.
%
As discussed in Section \ref{sec:static}, the static algorithm requires template directions to be specified by the user in order for the reachable set computation to proceed.
%
In practice, the parallelotopes tend to be defined by a combination of axis-aligned directions or diagonal directions.
%
However, it is unclear whether these template directions work reasonably well over general non-linear dynamics.
%
This is a problem which has profound consequences for the computation of useful over-approximations.
%
Choosing the incorrect directions could very well lead to an over-approximation too conservative for any practical use.
%
Figure \ref{fig:bloat_cvdp} portrays a simple example of a reachable set whose wrapping error becomes explosive after a few steps.
%
In light of these ramifications, it is esstentially incumbent on the user to choose fruitful template directions that control the growth of the over-approximation.
%
Due to the inherent difficult nature of choosing these template directions, it may be unreasonable to delegate this responsibility to a practitioner who knows little about parallelotope-based reachability.
%
These drawbacks to the static algorithm motivate the rationale behind the \emph{automatic} aspect of our modified algorithm.

Additionally, as the template directions are set at the beginning, these directions cannot adapt to the dynamics as the computation executes.
%
There may be template directions which may yield much leaner over-approximations at an intermediate step during the computation.
%
Hence, the ability to change the composition of the parallelotope bundle by adding and removing parallelotopes based on a data-driven approach could yield better over-approximations.
%
These considerations motivate the rationale behind the \emph{dynamic} aspect of our modified algorithm.

### 2022-08-19 thesis/prelim.tex
Formatting, font changes. Rearrangment of figures

\begin{equation}
[c_l(1), \;c_u(1)] \times [c_l(2),\; c_u(2)] \times \cdots \times [c_l(n), \; c_u(n)]
\end{equation}
%
It is worth noting that axis-aligned box on the 2D plane above would give the representation:
\begin{align}
  \Lambda = \begin{bmatrix*}[r]
            1 & 0 \\
            0 & 1 \\
            -1 & 0 \\
            0 & -1 \\
            \end{bmatrix*} , \quad c = [2,2,-1,-1]^T
\end{align}
if we were to convert the half-space representation above into the form defined in Equation \ref{eq:halfspace}.
%
From here on out, we refer to parallelotopes defined by the $n$ axis-aligned directions as the \emph{axis-aligned} parallelotopes.
%
For a visual plot of above axis-aligned parallelotope, see Figure \ref{fig:axis_align_ptope}.

\begin{figure}[h!]
  \centering
  \begin{tikzpicture}
    \draw[step=1cm,gray,very thin, style=help lines] (-3,-3) grid (3,3);
    \fill[blue!40!white, opacity=0.4] (-1,-1) rectangle (1,1);
    \draw[->, thick](0,1) -- (0,2) node[above] (0,2) {$[0,1]$};
    \draw[->, thick](1,0) -- (2,0) node[right] (2,0) {$[1,0]$};
    \draw[->, thick](-1,0) -- (-2,0) node[left] (-2,0) {$[-1,0]$};
    \draw[->, thick](0,-1) -- (0,-2) node[below] (0,-2) {$[0,-1]$};
    \draw[fill] (0,0) circle [radius=1pt];
    \node[above] (0,0) {$\mathcal{O}$};
    %\foreach \i in {-3,-2,...,3} \node[below] at (\i,0) {\scriptsize \i};
    %\foreach \i in {-3,-2,...,3} \node[right] at (0,\i) {\scriptsize \i};
  \end{tikzpicture}

  \caption{Plot of the axis-aligned parallelotope of Example \ref{ex:simple_ptope}. The template directions are displayed as normal vectors to the pairs of parallel planes defining the parallelotope. $\mathcal{O}$ represents the origin $(0,0)$ and each side of a grid cell represents one unit of distance.}
  \label{fig:axis_align_ptope}
\end{figure}

@@

\begin{example}
  \label{ex:simple_diag_ptope}
  We can also consider the axis-aligned directions rotated $45^\circ$ counter-clockwise. This would yield the two diagonal directions $[1,1]^T, [-1,1]^T$. Suppose we set the upper and lower offsets to be $c_u = [1,1]^T$ and $c_l = [-1,-1]^T$ respectively. Then once again by Definition \ref{def:halfspace_def}:
  %
  \begin{equation}
  \begin{bmatrix}
    -1 \\ -1
  \end{bmatrix} \leq \begin{bmatrix*}[r]
      1 & 1 \\ -1 & 1
    \end{bmatrix*}\begin{bmatrix}
     x \\ y
    \end{bmatrix} \leq \begin{bmatrix}
      1 \\ 1
    \end{bmatrix}
  \end{equation}
  %
  The bounded region in $\reals^2$ will be the conjunction of the linear inequalities:
  %
  \begin{align}
    \begin{split}
    -1 \leq x + y \leq 1 \\
    -1 \leq y - x \leq 1
    \end{split}
  \end{align}
  %
  In general, we define \emph{diagonal directions} are defined to be vectors created by adding and subtracting distinct pairs of unit axis-aligned vectors from each other. As a matter of convenience however, we refer to \emph{diagonal parallelotopes} as those defined by a combination of axis-aligned \emph{and} diagonal directions. This definition will be useful when we consider parallelotopes defined by unconventional directions (i.e those template directions which are neither axis-aligned nor diagonal). To see a visual plot of the diagonal parallelotope, see Figure \ref{fig:diagonal_ptope}.

  \begin{figure}[h!]
    \centering
    \begin{tikzpicture}
      \draw[step=1cm,gray,very thin, style=help lines] (-3,-3) grid (3,3);
      \draw[fill=blue!40!white, opacity=0.4] (1,0) -- (0,1) -- (-1,0) -- (0,-1) -- cycle;
      \draw[->, thick] (0.5,0.5) -- (1,1) node[above] {$[1,1]$};
      \draw[->, thick] (-0.5,-0.5) -- (-1,-1) node[below] {$[-1,-1]$};
      \draw[->, thick] (-0.5,0.5) -- (-1,1) node[above] {$[-1,1]$};
      \draw[->, thick] (0.5,-0.5) -- (1,-1) node[below] {$[1,-1]$};
      \draw[fill] (0,0) circle [radius=1pt];
      \node[above] {$\mathcal{O}$};
    \end{tikzpicture}
    \caption{Plot of the rotated parallelotope of Example \ref{ex:simple_diag_ptope}. Note the manner in which the template directions and their negative counterparts define parallel planes and the parallelotope's definition as the intersection of the postive and negative half-spaces of all pairs of parallel planes. $\mathcal{O}$ represents the origin $(0,0)$ and each side of a grid cell represents one unit of distance.}
    \label{fig:diagonal_ptope}
  \end{figure}
\end{example}

@@

As a final remark, notice that for a parallelotope $P$, the generator representation also defines an affine transformation that maps $[0,1]^{n}$ to $P$.
%
We refer to this affine transformation associated to $P$ as $T_P:[0,1]^n \rightarrow P$ when necessary.

>>>>

\begin{figure}[t!]
  \centering
  \begin{tikzpicture}
    \draw[step=1cm,gray,very thin, style=help lines] (-3,-3) grid (3,3);
    \fill[blue!40!white, opacity=0.4] (-1,-1) rectangle (1,1);
    \draw[->, thick](0,1) -- (0,2) node[above] (0,2) {$[0,1]$};
    \draw[->, thick](1,0) -- (2,0) node[right] (2,0) {$[1,0]$};
    \draw[->, thick](-1,0) -- (-2,0) node[left] (-2,0) {$[-1,0]$};
    \draw[->, thick](0,-1) -- (0,-2) node[below] (0,-2) {$[0,-1]$};
    \draw[fill] (0,0) circle [radius=1pt];
    \node[above] (0,0) {$\mathcal{O}$};
    %\foreach \i in {-3,-2,...,3} \node[below] at (\i,0) {\scriptsize \i};
    %\foreach \i in {-3,-2,...,3} \node[right] at (0,\i) {\scriptsize \i};
  \end{tikzpicture}

  \caption{Plot of the axis-aligned parallelotope of Example \ref{ex:simple_ptope}.}
  \label{fig:axis_align_ptope}
\end{figure}

\begin{equation}
[c_l(1), \;c_u(1)] \times [c_l(2),\; c_u(2)] \times \cdots \times [c_l(n), \; c_u(n)]
\end{equation}
%
It is worth noting that axis-aligned box on the 2D plane above would give the representation:
\begin{align}
  \Lambda = \begin{bmatrix*}[r]
            1 & 0 \\
            0 & 1 \\
            -1 & 0 \\
            0 & -1 \\
            \end{bmatrix*} , \quad c = [2,2,-1,-1]^T
\end{align}
if we were to convert the half-space representation above into the form defined in Equation \ref{eq:halfspace}.
%
From here on out, we refer to parallelotopes defined by the $n$ axis-aligned directions as the \emph{axis-aligned} parallelotopes.
%
For a visual plot of above axis-aligned parallelotope, see Figure \ref{fig:axis_align_ptope}. The template directions are displayed as normal vectors to the pairs of parallel planes defining the parallelotope. $\mathcal{O}$ represents the origin $(0,0)$ and each side of a grid cell represents one unit of distance.

@@

\begin{example}
  \label{ex:simple_diag_ptope}
  We can also consider the axis-aligned directions rotated $45^\circ$ counter-clockwise. This would yield the two diagonal directions $[1,1]^T, [-1,1]^T$. Suppose we set the upper and lower offsets to be $c_u = [1,1]^T$ and $c_l = [-1,-1]^T$ respectively. Then once again by Definition \ref{def:halfspace_def}:
  %
  \begin{equation}
  \begin{bmatrix}
    -1 \\ -1
  \end{bmatrix} \leq \begin{bmatrix*}[r]
      1 & 1 \\ -1 & 1
    \end{bmatrix*}\begin{bmatrix}
     x \\ y
    \end{bmatrix} \leq \begin{bmatrix}
      1 \\ 1
    \end{bmatrix}
  \end{equation}
  %
  The bounded region in $\reals^2$ will be the conjunction of the linear inequalities:
  %
  \begin{align}
    \begin{split}
    -1 \leq x + y \leq 1 \\
    -1 \leq y - x \leq 1
    \end{split}
  \end{align}
  %
  In general, we define \emph{diagonal directions} are defined to be vectors created by adding and subtracting distinct pairs of unit axis-aligned vectors from each other. As a matter of convenience however, we refer to \emph{diagonal parallelotopes} as those defined by a combination of axis-aligned \emph{and} diagonal directions. This definition will be useful when we consider parallelotopes defined by unconventional directions (i.e those template directions which are neither axis-aligned nor diagonal). To see a visual plot of the diagonal parallelotope, see Figure \ref{fig:diagonal_ptope}. Note the manner in which the template directions and their negative counterparts define parallel planes and the parallelotope's definition as the intersection of the postive and negative half-spaces of all pairs of parallel planes. Again, $\mathcal{O}$ represents the origin $(0,0)$ and each side of a grid cell represents one unit of distance.

  \begin{figure}[t!]
    \centering
    \begin{tikzpicture}
      \draw[step=1cm,gray,very thin, style=help lines] (-3,-3) grid (3,3);
      \draw[fill=blue!40!white, opacity=0.4] (1,0) -- (0,1) -- (-1,0) -- (0,-1) -- cycle;
      \draw[->, thick] (0.5,0.5) -- (1,1) node[above] {$[1,1]$};
      \draw[->, thick] (-0.5,-0.5) -- (-1,-1) node[below] {$[-1,-1]$};
      \draw[->, thick] (-0.5,0.5) -- (-1,1) node[above] {$[-1,1]$};
      \draw[->, thick] (0.5,-0.5) -- (1,-1) node[below] {$[1,-1]$};
      \draw[fill] (0,0) circle [radius=1pt];
      \node[above] {$\mathcal{O}$};
    \end{tikzpicture}
    \caption{Plot of the rotated parallelotope of Example \ref{ex:simple_diag_ptope}.}
    \label{fig:diagonal_ptope}
  \end{figure}
\end{example}

@@

\begin{remark}
Notice that for a parallelotope $P$, the generator representation also defines an affine transformation that maps $[0,1]^{n}$ to $P$.
%
We refer to this affine transformation associated to $P$ as $T_P:[0,1]^n \rightarrow P$ when necessary.
\end{remark}

### 2022-08-19 thesis/realtime.tex
thesis slide commit
Committing old format slides and more changes to thesis
Formatting, font changes. Rearrangment of figures

To mitigate this explosive growth in coefficients, we can pose the following question: Is there a method for ascertaining a small subset of the basis monomials such that the correct maximum and minimum coefficients can be computed from this subset rather the entire set of Bernstein basis monomials?
%
Towards this end, we propose the following algorithm outlined in Algorithm \ref{alg:bern_prune} and elaborate on the individual subroutines. In the context of these experiments, we assume the dynamics $f$ are polynomial. This is to ensure we have a Bernstein expansion of the composed dynamics $f \circ T_p$.

@@

\begin{enumerate}
\litem{Parition Domain into Grid Cells:} We first partition the domain into grid cells of equal dimension. Naturally the entire domain is infinite, so we must restrict the domain into a bounded box of sufficiently large dimension. This box must bound the reachable set computed for a finite number of steps forward. These grid cells will act as constraints to feed into our generator representation.

@@

@@

\begin{itemize}
  \item  We attempted to run the modified reachability algorithm for seven steps.
  \item We partitioned the domain and fixed the initial set for each model according to the below inputs:

  \begin{tabular}{|c|c|c|c|}
  \hline
  Model & Domain & Initial Set & \# of Grid Cells \\
  \hline
  Vanderpol & $[-3, 3] \times [-3, 3]$ & $[0.001, 0.005]\times [1.995, 2]$ & 1600 \\
  %\hline
  %SIR & $[0, 1]^3$ & $[0.799,0.8] \times [0.199,0.2] \times 0$ & 100 \\
  \hline
  \end{tabular}
\end{itemize}

@@

Consider the output presented in Figure \ref{fig:bern_prune_deg_output} computed with just the axis-aligned parallelotope over the VanderPol model with a grid parition of $40 \times 40$. We run the algorithm for one step and display the output during the optimization procedure for template direction $\Lambda_i = [0,1]^T$.
%-------------------------
\input{bern_prune_deg_output}
%-------------------------
During the first step, the maximum degree is a lone monomial:
%
\begin{verbatim}
Returned set of Maximum Monom from Lookup Table:
[((3, 3), 2.18359741950000)]
\end{verbatim}
%
This means that the actual monomial with the maximum Bernstein coefficient:
\begin{verbatim}
Actual Max Monom Deg: (2, 0)
\end{verbatim}
%
was pruned out during the pre-computation phase.
%
What seems to happen quite frequently is the maximum or minimum intervals and their overlaps dominate the degree of the polynomial computed after composition during the reachability computation. To see this, note the anchor and generator vectors computed for the generator representation:
%
\begin{verbatim}
BASE VERTEX: [0.005, 2.0]
GENERATORS: [[-0.004, 0.0], [0.0, -0.0049999999999999]]
\end{verbatim}

The generators are axis-aligned, meaning that several degrees are annihilated during the composition $\Lambda_i\cdot (f\circ T_P)$ due to the zero components.
%
Hence, during our pre-processing stage, we consider a more general set of monomial degrees which may not reflect the relevant degree statistics found during the actual computation phase.

This poses the issue of being privy to the degrees supplied to the algorithm before it even begins.
%
In other words, we are uncertain about how the interactions of the functional composition with the generator representation affect the degree of the resulting polynomial. If the degree is smaller than that of the monomial with the maximum/minimum coefficient interval or its overlaps, then it will appear during our reachability algorithm. This untowardly results in either a gross conservative error or an exception indicting that too many monomials were pruned out.
%
This error becomes further compounded if higher degree terms are wiped out by the zero factors in the computed generator or if the wrapping error becomes worse.

@@

As the wrapping error becomes worse, so does the utility of our lookup table. A guess is that as the conservativeness of our reachable becomes greater, it overlaps with more cells of the domain. This results in an cascading effect where the reachability algorithm is required to take into account more monomials as it progresses.

We speculate that the utility of our pruning method hinges on the careful control of the reachable set error. If the error becomes too great, we end of computing all of the coefficients, negating any benefits of a speed-up.

>>>>

To mitigate this explosive growth in coefficients, we pose the following question:
%
\begin{tcolorbox}[standard jigsaw, opacityback=0]
Does there exists a method for ascertaining a small subset of the basis monomials such that the correct maximum and minimum coefficients can be computed from this subset rather the entire set of Bernstein basis monomials?
\end{tcolorbox}
%
To this end, we propose the following algorithm outlined in Algorithm \ref{alg:bern_prune} and elaborate on the individual subroutines. In the context of these experiments, we assume the dynamics $f$ are polynomial. This is to ensure we have a Bernstein expansion of the composed dynamics $f \circ T_p$.

@@

\begin{enumerate}[leftmargin=*]
\litem{Parition Domain into Grid Cells:} We first partition the domain into grid cells of equal dimension. Naturally the entire domain is infinite, so we must restrict the domain into a bounded box of sufficiently large dimension. This box must bound the reachable set computed for a finite number of steps forward. These grid cells will act as constraints to feed into our generator representation.

@@

\input{mod_reach_alg}

@@

\begin{itemize}
  \item  We attempted to run the modified reachability algorithm for seven steps.
  \item We partitioned the domain and fixed the initial set for each model according to the parameters listed in Table~\ref{tab:vdp_rt}.
\end{itemize}

\begin{table}
 \centering
 \begin{tabular}{|c|c|c|c|}
 \hline
 Model & Domain & Initial Set & \# of Grid Cells \\
 \hline
 Vanderpol & $[-3, 3] \times [-3, 3]$ & $[0.001, 0.005]\times [1.995, 2]$ & 1600 \\
 %\hline
 %SIR & $[0, 1]^3$ & $[0.799,0.8] \times [0.199,0.2] \times 0$ & 100 \\
 \hline
 \end{tabular}
 \caption{Parameters for Vanderpol Model used during Real-time Reachability Experiments}
 \label{tab:vdp_rt}
\end{table}

@@

Consider the output presented in Figure \ref{fig:bern_prune_deg_output} computed with just the axis-aligned parallelotope over the VanderPol model with a grid parition of $40 \times 40$. We run the algorithm for one step and display the output during the optimization procedure for template direction $\Lambda_i = [0,1]^T$.
%-------------------------
\input{bern_prune_deg_output}
%-------------------------
During the first step, the maximum degree is a lone monomial:
%
\begin{verbatim}
Returned set of Maximum Monom from Lookup Table:
[((3, 3), 2.18359741950000)]
\end{verbatim}
%
This means that the actual monomial with the maximum Bernstein coefficient:
\begin{verbatim}
Actual Max Monom Deg: (2, 0)
\end{verbatim}
%
was pruned out during the pre-computation phase.
%
Frequently, the maximum or minimum intervals and their overlaps dominate the degree of the polynomial computed after composition during the reachability computation. To see this, note the anchor and generator vectors computed for the generator representation:
%
\begin{verbatim}
BASE VERTEX: [0.005, 2.0]
GENERATORS: [[-0.004, 0.0], [0.0, -0.0049999999999999]]
\end{verbatim}

The generators are axis-aligned, meaning that several degrees are annihilated during the composition $\Lambda_i\cdot (f\circ T_P)$ due to the zero components.
%
Hence, during our pre-processing stage, we consider a more general set of monomial degrees which may not reflect the relevant degree statistics found during the actual computation phase.
%
This poses the issue of being privy to the degrees supplied to algorithm before execution begins.
%
In other words, we are uncertain about how the interactions of the functional composition with the generator representation affect the degree of the resulting polynomial. If the degree is smaller than that of the monomial with the maximum/minimum coefficient interval or its overlaps, then it will appear during our reachability algorithm. This untowardly results in either a gross conservative error or an exception indicting that too many monomials were pruned out.
%
This error becomes further compounded if higher degree terms are wiped out by the zero factors in the computed generator or if the wrapping error becomes worse.

@@

As the wrapping error becomes worse, so does the utility of our lookup table. A guess is that as the conservativeness of our reachable becomes greater, it overlaps with more cells of the domain. This results in a cascading effect where the reachability algorithm is required to take into account more monomials as it progresses.
%
In light of the discussed observations, we speculate that the utility of our pruning method hinges on the careful control of the reachable set error. If the error becomes too great, we end of computing all of the coefficients, negating any benefits of a speed-up.
